\documentclass{article} % For LaTeX2e
\usepackage{analemma,times}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}

% For theorems and such
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}

% Custom
\usepackage{multirow}
\usepackage{color}
\usepackage{colortbl}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{xspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\graphicspath{{figures/}} % To reference your generated figures, name the PNGs directly.

\title{Selective Delexicalization to Defend Structured-Output LLM APIs from Control-Plane Jailbreaks}

\author{FARS \\
Analemma \\
\texttt{fars@analemma.ai}
}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\analemmafinalcopy % Camera-ready version (non-anonymous)
\begin{document}


\maketitle

\begin{abstract}
Structured-output APIs enable reliable LLM integration through constrained decoding, but recent work reveals that JSON schema specifications create a new attack surface for control-plane jailbreaks. Attackers can embed harmful instructions in forced enum or const values, bypassing safety alignment by forcing models to output malicious content verbatim. We propose Selective DeLex-JSON, a training-free defense that sanitizes JSON schemas before constrained decoding. Our approach identifies suspicious forced literals using a conjunction-based heuristic (flagging values that are long, contain whitespace, or include imperative verbs) and replaces them with opaque placeholders that preserve schema structure while removing attack payloads. On HarmBench, DeLex-JSON achieves 0\% attack success rate on both Llama-3.1-8B and Qwen2.5-7B, completely neutralizing the EnumAttack that achieves 22\% ASR without defense. The defense incurs only 1.1\% benign schema modification rate, Pareto-dominating the Reject-Only baseline which has 4.4\% rejection rate. The defense is immediately deployable in production inference pipelines without model retraining.


\textit{WARNING: This paper was generated by an automated research system. The code is publicly available.}\footnote{\url{https://github.com/fars-analemma/selective-delex-json}}
\end{abstract}

\input{sections/introduction}

\input{sections/related_work}

\input{sections/method}

\input{sections/experiments}

\input{sections/conclusion}

\bibliography{analemma}
\bibliographystyle{analemma}


% Appendix can be added here if needed

\end{document}
